---
title: "Plato's Pizza"
author: "Remy Drain"
date: "Last updated `r Sys.Date()`"
output: 
  html_notebook:
    toc: true
bibliography: sources.bib
csl: chicago-fullnote-bibliography.csl
suppress-bibliography: true
---

## Process
### Load Packages
```{r install and load packages, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(RSQLite)
library(forcats)
library(ggplot2)
```

### Read in the datasets
```{r read in data, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
dict <- read.csv(file.path(getwd(), "sales-data/data_dictionary.csv"))
order_details <- read.csv(file.path(getwd(), "sales-data/pizza_sales/order_details.csv"))
orders <- read.csv(file.path(getwd(), "sales-data/pizza_sales/orders.csv"))
pizza_types <- read.csv(file.path(getwd(), "sales-data/pizza_sales/pizza_types.csv"))
pizzas <- read.csv(file.path(getwd(), "sales-data/pizza_sales/pizzas.csv"))
```

### Explore the data
I've got my data imported, including the data dictionary that was provided. I'll use tibbles to get an idea of what the data looks like in each of the tables.
```{r}
tibble(dict)
tibble(order_details)
tibble(orders)
tibble(pizza_types)
tibble(pizzas)
```

I want to aggregate some of the data and create at least one new variable showing the day of the week, so I'm going to use the `RSQLite` package to create a virtual database that I can run SQL queries through, then save them as dataframes for plotting.

### Create a SQLite Database
```{r create database, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
db <- dbConnect(RSQLite::SQLite(), dbname=":memory:")

dbWriteTable(db, "order_details", order_details)
dbWriteTable(db, "orders", orders)
dbWriteTable(db, "pizza_types", pizza_types)
dbWriteTable(db, "pizzas", pizzas)

dbListTables(db)
```

I'll start by creating a dataframe for the orders, where I want to include the following:

* New column for weekday[^1]
* Order total

```{sql add weekdays variable, connection=db, output.var=odf}
SELECT
  orders.order_id,
  strftime('%w', julianday(date)) AS weekday,
  date,
  time,
  SUM(order_details.quantity) AS total_items,
  SUM(pizza_price) AS order_total
FROM orders
JOIN
  (SELECT
    order_details.order_id,
    order_details.quantity,
    pizzas.pizza_id,
    pizzas.price*order_details.quantity AS pizza_price
  FROM order_details
  JOIN pizzas
    ON pizzas.pizza_id=order_details.pizza_id) AS order_details
  ON order_details.order_id=orders.order_id
GROUP BY orders.order_id
```

Since SQLite doesn't have a data type for date and time,[^2] I'll want to convert some of the columns in our dataframe so I can use them for plotting.

I also want to change the values in the weekday column to the actual days of the week, so I'll use R to replace them with their corresponding values.[^3]
```{r convert weekday values}
odf['weekday'][odf['weekday'] == '0'] <- 'Sunday'
odf['weekday'][odf['weekday'] == '1'] <- 'Monday'
odf['weekday'][odf['weekday'] == '2'] <- 'Tuesday'
odf['weekday'][odf['weekday'] == '3'] <- 'Wednesday'
odf['weekday'][odf['weekday'] == '4'] <- 'Thursday'
odf['weekday'][odf['weekday'] == '5'] <- 'Friday'
odf['weekday'][odf['weekday'] == '6'] <- 'Saturday'
```

Et, voilá! I've got my first plottable dataframe. Let's take it for a spin with `ggplot2`.
```{r plotting orders}

```


[^1]: <https://www.sqlite.org/lang_datefunc.html#strftm>
[^2]: <https://www.sqlite.org/datatype3.html>
[^3]: <https://www.statology.org/replace-values-in-data-frame-r/>