---
title: "Plato's Pizza Analysis"
author: "Remy Drain"
date: "Last updated: `r Sys.Date()`"
output: 
  html_notebook:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```
## Introduction
Plato's Pizza is a Greek-inspired pizza place in New Jersey looking for ways to improve operations using data they've collected over the last year. They've asked us to analyze the data and put together a report to help them find opportunities to drive more sales and work more efficiently.[^1]

## Problems
Some specific questions they've asked:

* What days and times do we tend to be busiest?
* How many pizzas are we making during peak periods?
* What are our best and worst selling pizzas?
* What's our average order value?
* How well are we using our seating capacity? (we have 15 tables and 60 seats)

## Data
The data has been provided in the following 5 files:

* **Data Dictionary:** Contains a breakdown of the variables contained in each table, along with a brief description
* **Order Details:** Contains details of the orders, and connects with the orders table on order_id and the pizzas table on pizza_id
* **Orders:** Contains the id, date, and time of the order
* **Pizza Types:** Contains the name, category, and ingredients of each pizza, and connects with the pizzas table on pizza_type_id
* **Pizzas:** Contains the id, size, and price of the pizzas

## Process
### Getting to Know the Data
I imported the datasets into BigQuery to perform some aggregation and analysis using SQL.

### Read in datasets
```{r read in data, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
dict <- read.csv(file.path(getwd(), "sales-data/data_dictionary.csv"))
order_details <- read.csv(file.path(getwd(), "sales-data/pizza_sales/order_details.csv"))
orders <- read.csv(file.path(getwd(), "sales-data/pizza_sales/orders.csv"))
pizza_types <- read.csv(file.path(getwd(), "sales-data/pizza_sales/pizza_types.csv"))
pizzas <- read.csv(file.path(getwd(), "sales-data/pizza_sales/pizzas.csv"))
odf <- read.csv(file.path(getwd(), "sales-data/agg_orders.csv"))
```


[^1]: <https://mavenanalytics.io/challenges/maven-pizza-challenge/4>